HO unification from object language to meta language

### 2024-02-06 main.tex
tex add example of FO and HO unification problems from stdpp

\begin{mintedTcolorbox}{coq}
  Class XXX Args := {methods ...}.
  Instance ZZZ : bla bla -> XXX Args := {...} Qed.
  Instance WWW : bla bla -> XXX Args := {...} Qed.
\end{mintedTcolorbox}

@@

\begin{mintedTcolorbox}{elpi}
  type XXX Args .. -> term -> o.
  XXX Args ... (WWW P1 ... PN) :-
    premese...
  XXX Args ... (ZZZ P1 ... PN) :-
    premese...
\end{mintedTcolorbox}

Type-class resolution starts from a query, that is a class applied to some
argument. This \coq term is translated into a term of the meta language and the
search for a solution in the database is started. However, since unification is
performed on the head of the clause, it may happen that the term representation
in the meta language may hide some unification properties that are true in the
object language.

>>>>

\begin{mintedTcolorbox}[escapeinside=&&]{coq}
  Inductive sig (A : Type) (P : A -> Prop) : Type := ...

  Class Decision (P : Prop) := decide : {P} + {not P}.
  Class RelDecision {A B: Type} (R : A -> B -> Prop).
  Class ProofIrrel (A : Type) : Prop := proof_irrel (x y : A) : x = y.

  Instance decide_rel: forall (A B : Type) (R : A -> B -> Prop),
    RelDecision R -> forall (x : A) (y : B), Decision (R x y). Admitted.
  Instance True_pi : ProofIrrel True. Admitted.
  Instance sig_eq_dec: forall (A : Type) (P : A -> Prop),
    (forall x, ProofIrrel (P x)) -> RelDecision (@eq A) ->
    RelDecision (@eq (sig A P)). Admitted.
\end{mintedTcolorbox}

@@

\begin{mintedTcolorbox}{elpi}
  type tc-Decision term -> term -> o.
  type tc-RelDecision term -> term -> term -> term -> o.
  type tc-ProofIrrel term -> term -> o.

  tc-ProofIrrel (`True`) (`True_pi`).
  tc-Decision (app [R, X, Y]) 
    (app [`decide_rel`, A, B, R, P, X, Y]) :-
    tc-RelDecision A B R P.
  tc-RelDecision (app [`sig`, A, P]) 
    (app [`sig`, A, P]) 
    (app [`eq`, app [`sig`, A, P]]) 
    (app [`sig_eq_dec`, A, P, P1, P2]) :-
    pi-decl c0 `x` A =>
      tc-ProofIrrel (app [P, c0]) (app [P1, c0]), 
    tc-RelDecision A A (app [`eq`, A]) P2.
\end{mintedTcolorbox}

Type-class resolution starts from a query, that is a class applied to some
argument. This \coq term is translated into a term of the meta language and the
search for a solution in the database is started. However, since unification is
performed on the head of the clause, it may happen that the term representation
in the meta language may hide some unification properties that are true in the
object language. In the example below, the goal \texttt{Decision (@eq T a b)}
for some \texttt{a} and \texttt{b} unifies with \texttt{Decision (R x y)} in the
object language but not in its meta representation. Similarly, the goal
\texttt{RelDecision (@eq (sig T ?P))} where \texttt{?P}, under the hypothesis
\texttt{RelDecision (@eq nat)}, will trying to apply the rule for
\texttt{sig\_eq\_dec}, we fall into an higher order unification problem, where
\texttt{P} is applied to the local name \texttt{x}. However, the corresponding
rule in the meta language exploit a first order variable \texttt{P}. Therefore,
after the refinement of the goal to \texttt{sig\_eq\_dec}, the resolution
immediately fail to solve the premise \texttt{tc-ProofIrrel (app [P, c0]) (app
[P1, c0])}.
